Switchboard and BNC Suprisal

### 2019-09-17 surprisal/bnc&switchboard.Rmd
begin implementation of Kneser-Ney smoothing

switch_trigram_surprisals <- map_dfr(1:nrow(sts),
        ~bind_cols(select(slice(sts, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(sts, .x)),
                                             column = s, nboot = 1000)))
# switch_trigram_surprisals <- trigram_surprisals_prep %>%
#   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
#                     ifelse(is.na(lag_word1), # check if not the second word
#                            -log(p), # trigram if third or beyond
#                            -log(cond_p)), # bigram if second word
#                     -log(tri_cond_p))) %>% # unigram if first word
#   group_by(length, word_order) %>%
#   summarise(s = mean(s)) %>% 
#   filter(word_order <= length)

@@

```{r bnc}
bnc <- read_feather(here("../bnc/bnc10.feather")) %>%
  bind_rows(read_feather(here("../bnc/bnc20.feather"))) %>%
  bind_rows(read_feather(here("../bnc/bnc30.feather")))
```

@@

bnc_surprisals <- bnc %>%
  left_join(bnc_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

@@

bnc_surprisals <- map_dfr(1:nrow(bnc_surprisals_prep), 
        ~bind_cols(select(slice(bnc_surprisals_prep, .x), length, word_order), 
                               tidyboot_mean(data = unnest(slice(bnc_surprisals_prep, .x)), 
                                             column = s, nboot = 100)))

# bnc_surprisals <- bnc %>%
#   left_join(bnc_unigrams) %>%
#   mutate(s = -log(p)) %>%
#   select(length, word_order, s) %>%
#   filter(length == 10) %>%
#   group_by(length, word_order) %>%
#   # summarise(s = mean(s)) 
#   tidyboot_mean(s)

@@

```{r bnc bigrams}
bnc_bigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  mutate(joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(bnc_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = joint_p / p) %>%
  select(-n, -p)

bigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(bnc_bigrams) %>%
  left_join(bnc_unigrams)

bnc_bigram_surprisals <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)
```

```{r bnc trigrams}
bnc_trigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(bnc_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-joint_p, -tri_joint_p, -cond_p) # %>%

trigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(bnc_trigrams) %>% 
  left_join(select(bnc_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(bnc_unigrams)

bnc_trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

bnc_trigram_surprisals %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

>>>>

switch_trigram_surprisals <- map_dfr(1:nrow(sts),
        ~bind_cols(select(slice(sts, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(sts, .x)),
                                             column = s, nboot = 1000)))
```

@@

```{r bnc}
bnc <- read_feather(here("../bnc/bnc10.feather")) %>%
  bind_rows(read_feather(here("../bnc/bnc20.feather"))) %>%
  bind_rows(read_feather(here("../bnc/bnc30.feather")))

selected_bnc <- bnc %>% 
  select(length, utterance_id) %>%
  distinct() %>%
  group_by(length) %>%
  sample_n(8000) %>% # downsampling because otherwise there are at least 70000 utterances for each length
  ungroup() %>%
  select(utterance_id)

bnc <- bnc %>%
  semi_join(selected_bnc)
```

@@

@@

bnc_surprisals <- map_dfr(1:nrow(bnc_surprisals_prep), 
        ~bind_cols(select(slice(bnc_surprisals_prep, .x), length, word_order), 
                               tidyboot_mean(data = unnest(slice(bnc_surprisals_prep, .x)), 
                                             column = s, nboot = 1000)))  
```

```{r bnc bigrams}
bnc_bigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(bnc_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x / n.y) %>%
  select(-n.y, -p) %>%
  rename(n = n.x)

bigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  left_join(bnc_bigrams, by = c("word", "lag_word")) %>%
  left_join(bnc_unigrams, by = "word")

bbs <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s)

bnc_bigram_surprisals <- map_dfr(1:nrow(bbs),
        ~bind_cols(select(slice(bbs, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(bbs, .x)),
                                             column = s, nboot = 1000)))

@@

```{r bnc trigrams}
bnc_trigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(bnc_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n.x / n.y) %>%
  select(-n.x, -n.y, -cond_p)

trigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(bnc_trigrams) %>% 
  left_join(bnc_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(bnc_unigrams, by = c("word"))

bts <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s)

bnc_trigram_surprisals <- map_dfr(1:nrow(bts),
        ~bind_cols(select(slice(bts, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(bts, .x)),
                                             column = s, nboot = 1000)))

```{r saving switchboard surprisals}
bnc_surprisals %>% 
  mutate(context = 0, corpus = "BNC") %>%
  bind_rows(bnc_bigram_surprisals %>% mutate(context = 1, corpus = "BNC")) %>% 
  bind_rows(bnc_trigram_surprisals %>% mutate(context = 2, corpus = "BNC")) %>% 
  write_feather(here("Data/bnc_bootstrap.feather"))
```
